---
title: "Untitled"
author: "Angie Boysen"
date: "July 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyr)
require(graphics); require(grDevices)
library(Hmisc)
library(gtools)
library(cowplot)
require(RColorBrewer)
library(xlsx)
library(readr)
library(plotly)
library(stringr)
library(GGally)
library(dplyr)
```

##Exploring the HILIC bacteria samples

#Import data

```{r echo=FALSE}
mydata.overloaded <- read_csv("./QC_output_Bacteria_HILIC_160718_1to100_AKB.csv")
mydata.overloaded <-mydata.overloaded[,2:11]
mydata <- mydata.overloaded
glimpse(mydata)
```

##Rename Runs
```{r echo=FALSE}
mydata$Replicate.Name[which(mydata$Replicate.Name == "160613_Blk_6-13 Blank")] <- "160613_Blk_FltBlk-0613_1to100"
mydata$Replicate.Name[which(mydata$Replicate.Name == "160613_Poo_6-13 Aq Extracts_1to100_1")] <- "160613_Poo_Poo-0613_1to100_1"
mydata$Replicate.Name[which(mydata$Replicate.Name == "160613_Poo_6-13 Aq Extracts_1to100_2")] <- "160613_Poo_Poo-0613_1to100_2"
mydata$Replicate.Name[which(mydata$Replicate.Name == "160613_Poo_6-13 Aq Extracts_1to100_3")] <- "160613_Poo_Poo-0613_1to100_3"

mydata <- mydata %>% mutate(Replicate.Name = Replicate.Name %>%
                                str_replace("Aq_",""))
mydata <- mydata %>% mutate(Replicate.Name = Replicate.Name %>%
                                str_replace("-gly","gly"))
mydata
```


# Here's an interactive plot
```{r eval=FALSE}
p <- ggplot(mydata, aes(x=Retention.Time, y=Area, colour = Compound.Name)) + geom_point(aes(text = paste(Replicate.Name)), alpha = 0.4) + scale_y_log10()
ggplotly(p)
```


##Read in Internal Standard list
```{r}
IS.list <- read_csv("M:/Summer_2016_metabolites/Methods_Paper_Samples/HILIC_MasterList_Summer2016.csv")  
IS.cat <- IS.list %>% filter(Group == "Internal Std") %>% 
     select(`Compound Name`) %>% 
     unique()
```


##Normalize to each internal Standard for which there aren't NAs
```{r}
wArea<- mydata %>%
     select(Replicate.Name,Compound.Name,Area) %>%
     spread(key=Compound.Name, value=Area) %>%
     as.data.frame

this.IS <- IS.cat[1,]
area.norm <- wArea[,-1] %>% 
     sapply(FUN = function(x) x/wArea[,grep(this.IS,
                                            names(wArea))]) %>%
     as_data_frame %>% 
     mutate(Replicate.Name = wArea$Replicate.Name) %>%
     gather(Compound,Area_Norm, -Replicate.Name)
key <- ncol(area.norm)
count <- length(which(!is.na(area.norm$Area_Norm)))/
               length(unique(area.norm$Replicate.Name))
names(area.norm)[key] <- paste(this.IS,"Norm.Area",sep=".")
print(paste(1, this.IS, count, sep="-"))

for (i in 2:nrow(IS.cat)){
     this.IS <- IS.cat[i,]
     if(length(wArea[,grep(this.IS, names(wArea))])!=0){
          this.norm <- wArea[,-1] %>% 
               sapply(FUN = function(x) x/wArea[,grep(this.IS,
                                                      names(wArea))]) %>%
               as_data_frame %>% mutate(Replicate.Name = wArea$Replicate.Name) %>%
               gather(Compound,Area_Norm, -Replicate.Name)
          key <- ncol(area.norm)
          area.norm[,key+1] <- this.norm$Area_Norm
          names(area.norm)[key+1] <- paste(this.IS,"Norm.Area",sep=".")
          count <- length(which(!is.na(this.norm$Area_Norm)))/
               length(unique(this.norm$Replicate.Name))
          print(paste(i, this.IS, count, sep="-"))
     }
}
```

#Normalize to Internal Standard
```{r echo=FALSE, eval=FALSE}
wArea<- mydata %>%
     select(Replicate.Name,Compound.Name,Area) %>%
     spread(key=Compound.Name, value=Area) %>%
     as.data.frame
wArea.norm <- wArea[,-1] %>% 
     sapply(FUN = function(x) x/wArea[,grep("Heavy Histidine",names(wArea))]) %>%
     as_data_frame %>% mutate(Replicate.Name = wArea$Replicate.Name) %>%
     gather(Compound,Area_Norm, -Replicate.Name)
wArea.norm
```


#Break Up the Names
Name structure must be:
Date_type_ID.duplicateID_dilution_injectionReplicate#

```{r echo=FALSE}
mydata_new <- area.norm %>% separate(Replicate.Name, 
                                      c("runDate",
                                        "type","ID.duplicate",
                                        "dilution","injReplicate"),"_") %>%
     mutate(Run.Cmpd = 
                 paste(area.norm$Replicate.Name,area.norm$Compound))
mydata <- mydata %>% mutate(Run.Cmpd = paste(Replicate.Name, Compound.Name))

dat <- full_join(mydata, mydata_new)
dat <- dat %>% select(-Compound,-rawArea)
glimpse(dat)
```

##Check the similarity of the pooled runs
```{r echo=FALSE}
pool.dat <- dat %>% filter(type == "Poo")
pool.dat <- pool.dat[-(grep("overload",x=pool.dat$Notes)),]


Norm.Names <- colnames(pool.dat)
Norm.Names

pool.stats <-  pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(Area_Norm, na.rm=T), 
                sd = sd(Area_Norm, na.rm=T), rsd = sd/m)

p <- ggplot(pool.stats, aes(x=rsd, text = Compound.Name)) + geom_histogram(binwidth = 0.05) + ggtitle("Relative Standard Deviation for 54 compounds in the pooled samples")
ggplotly(p)

wide.pool.dat <- pool.dat %>% 
     select(Compound.Name, injReplicate, Area_Norm) %>% 
     spread(key = injReplicate, value = Area_Norm)


p4<- ggplot(wide.pool.dat, aes(x=`1`, xend=`2`, 
                          y=`3`, yend=`3` , 
                          text=Compound.Name)) + 
     geom_segment() + geom_abline(slope = 1, intercept = 0) + 
     scale_x_log10() + scale_y_log10() + 
     xlab("Pooled replicates 1 and 2") + 
     ylab("Pooled replicate 3") + 
     ggtitle("Replicate injections of the pooled sample")
ggplotly(p4)
```

##Compare normalizations
```{r}
pool.dat <- dat %>% filter(type == "Poo")
pool.dat <- pool.dat[-(grep("overload",x=pool.dat$Notes)),]

sums <- pool.dat %>% select(Replicate.Name,Compound.Name,16:34) %>%
     gather(IS.Compound, NormArea, -Replicate.Name, -Compound.Name) %>%
     group_by(IS.Compound) %>% summarize(sums=sum(NormArea, na.rm=T))
Norm.Names <- sums %>% filter(sums > 0) %>% select(IS.Compound)
Norm.Names

##Initial normalization
##13C Acetyl CoA.Norm.Area
pool.stats <-  pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`13C Acetyl CoA.Norm.Area`, na.rm=T), 
                sd = sd(`13C Acetyl CoA.Norm.Area`, na.rm=T), rsd = sd/m)
rsd.dat <- pool.stats %>% select(Compound.Name) %>%
     mutate(`13C Acetyl CoA.Norm.Area` = pool.stats$rsd)

##13C-Isethionic Acid.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`13C-Isethionic Acid.Norm.Area`, na.rm=T), 
               sd = sd(`13C-Isethionic Acid.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`13C-Isethionic Acid.Norm.Area` =
                                   pool.stats$rsd)
##13C-Sulfoacetic Acid.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`13C-Sulfoacetic Acid.Norm.Area`, na.rm=T), 
               sd = sd(`13C-Sulfoacetic Acid.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`13C-Sulfoacetic Acid.Norm.Area` =
                                   pool.stats$rsd)

##`13C-Sulfolactic Acid.Norm.Area`
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`13C-Sulfolactic Acid.Norm.Area`, na.rm=T), 
               sd = sd(`13C-Sulfolactic Acid.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`13C-Sulfolactic Acid.Norm.Area` =
                                   pool.stats$rsd)

##Heavy Histidine.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`Heavy Histidine.Norm.Area`, na.rm=T), 
               sd = sd(`Heavy Histidine.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`Heavy Histidine.Norm.Area` =
                                   pool.stats$rsd)

##Heavy Isoleucine.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`Heavy Isoleucine.Norm.Area`, na.rm=T), 
               sd = sd(`Heavy Isoleucine.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`Heavy Isoleucine.Norm.Area` =
                                   pool.stats$rsd)


```


##make a figure comparing the normalizations
```{r norm.fig, echo=FALSE}
long.rsd <- rsd.dat %>% gather(Internal.Std, rsd, -Compound.Name)
sub <- long.rsd %>% filter(rsd < 0.99)
ggplot(sub, aes(x=Compound.Name, y=rsd, 
                shape=Internal.Std, color=Internal.Std)) +
     geom_point()  + ylim(0.5,0) +
     scale_shape_manual(values=c(19,19,19,17,17,17,17)) +
     scale_colour_manual(values=c("#e41a1c","#377eb8","#4daf4a",
                                  "#377eb8","#4daf4a","#984ea3",
                                  "#ff7f00")) + 
     theme_linedraw()+ 
     theme(axis.text.x = element_text(angle = 90, 
                                      hjust = 1, vjust = 0.5))

```

## Save out rsd data
```{r}
write.csv(long.rsd, file = "HILIC_1to100_RSD_NormalizationComparison_pool0613.csv")

write.csv(dat, file = "HILIC_1to100_data_NormalizationComparison.csv")

```


##Normalize to optical density
Incomplete
```{r echo=FALSE, eval=FALSE}
dens <- read.xlsx("M:/Summer_2016_metabolites/Methods_Paper_Samples/Bacteria_Skyline_Samples/Cultures_Extraction_Log.xlsx", 1)  # read first sheet

dens.order <- order(dens$Vial.ID)[1:16]

new.df$OD <- as.numeric(as.character(dens$OD.1.10..at.Harvest.time[dens.order]))

OD.Norm.df <- apply(new.df,2,function(x) x/new.df[,"OD"])
```




