---
title: "Untitled"
author: "Angie Boysen"
date: "July 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyr)
require(graphics); require(grDevices)
library(Hmisc)
library(gtools)
library(cowplot)
require(RColorBrewer)
library(xlsx)
library(readr)
library(plotly)
library(stringr)
library(GGally)
library(dplyr)
```

##Exploring the HILIC bacteria samples

#Import data

```{r echo=FALSE}
mydata.overloaded <- read_csv("./QC.output.new.Bacteria_HILIC_160613_1to2_update_akb.csv")
mydata.overloaded <-mydata.overloaded[,2:11]
mydata <- mydata.overloaded
glimpse(mydata)
```

##Rename Runs
```{r echo=FALSE}
mydata$Replicate.Name[which(mydata$Replicate.Name == 
                                  "160613_Blk_6-13 Blank")] <-
     "160613_Blk_FltBlk-0613_1to2"

mydata <- mydata %>% mutate(Replicate.Name = Replicate.Name %>%
                                str_replace("_Aq-Extracts","")) %>%
     mutate(Replicate.Name = Replicate.Name %>%
                                str_replace("6-","Poo-06"))

mydata <- mydata %>% mutate(Replicate.Name = Replicate.Name %>%
                                str_replace("Aq_",""))
mydata <- mydata %>% mutate(Replicate.Name = Replicate.Name %>%
                                str_replace("-gly","gly"))
mydata
```

##Read in Internal Standard list
```{r}
IS.list <- read_csv("M:/Summer_2016_metabolites/Methods_Paper_Samples/HILIC_MasterList_Summer2016.csv")  
IS.cat <- IS.list %>% filter(Group == "Internal Std") %>% 
     select(`Compound Name`) %>% 
     unique()
```

##Normalize to each internal Standard for which there aren't NAs
```{r}
wArea<- mydata %>%
     select(Replicate.Name,Compound.Name,Area) %>%
     spread(key=Compound.Name, value=Area) %>%
     as.data.frame

this.IS <- IS.cat[1,]
area.norm <- wArea[,-1] %>% 
          sapply(FUN = function(x) x/wArea[,grep(this.IS,
                                                 names(wArea))]) %>%
          as_data_frame %>% mutate(Replicate.Name = wArea$Replicate.Name) %>%
          gather(Compound,Area_Norm, -Replicate.Name)
key <- ncol(area.norm)
count <- length(which(!is.na(this.norm$Area_Norm)))/
               length(unique(this.norm$Replicate.Name))
names(area.norm)[key] <- paste(this.IS,"Norm.Area",sep=".")
print(paste(1, this.IS, count, sep="-"))

for (i in 2:nrow(IS.cat)){
     this.IS <- IS.cat[i,]
     if(length(wArea[,grep(this.IS, names(wArea))])!=0){
          this.norm <- wArea[,-1] %>% 
               sapply(FUN = function(x) x/wArea[,grep(this.IS,
                                                      names(wArea))]) %>%
               as_data_frame %>% mutate(Replicate.Name = wArea$Replicate.Name) %>%
               gather(Compound,Area_Norm, -Replicate.Name)
          key <- ncol(area.norm)
          area.norm[,key+1] <- this.norm$Area_Norm
          names(area.norm)[key+1] <- paste(this.IS,"Norm.Area",sep=".")
          count <- length(which(!is.na(this.norm$Area_Norm)))/
               length(unique(this.norm$Replicate.Name))
          print(paste(i, this.IS, count, sep="-"))
     }
}
```


#Normalize to Internal Standard
```{r echo=FALSE, eval=FALSE}
wArea<- mydata %>%
     select(Replicate.Name,Compound.Name,Area) %>%
     spread(key=Compound.Name, value=Area) %>%
     as.data.frame
wArea.norm <- wArea[,-1] %>% 
     sapply(FUN = function(x) x/wArea[,grep("Heavy Histidine",names(wArea))]) %>%
     as_data_frame %>% mutate(Replicate.Name = wArea$Replicate.Name) %>%
     gather(Compound,Area_Norm, -Replicate.Name)
wArea.norm
```


#Break Up the Names
Name structure must be:
Date_type_ID.duplicateID_dilution_injectionReplicate#

```{r echo=FALSE}
mydata_new <- area.norm %>% separate(Replicate.Name, 
                                      c("runDate",
                                        "type","ID.duplicate",
                                        "dilution","injReplicate"),"_") %>%
     mutate(Run.Cmpd = paste(area.norm$Replicate.Name,area.norm$Compound))
mydata <- mydata %>% mutate(Run.Cmpd = paste(Replicate.Name, Compound.Name))

dat <- full_join(mydata, mydata_new)
dat <- dat %>% select(-Compound)
glimpse(dat)
```

##Check the similarity of the pooled runs
```{r echo=FALSE}
pool.dat <- dat %>% filter(type == "Poo") %>% filter(ID.duplicate == "Poo-0614")
pool.dat <- pool.dat[-(grep("overload",x=pool.dat$Notes)),]

Norm.Names <- colnames(pool.dat)
Norm.Names

pool.stats <-  pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`13C Acetyl CoA.Norm.Area`, na.rm=T), 
                sd = sd(`13C Acetyl CoA.Norm.Area`, na.rm=T), rsd = sd/m)
rsd.dat <- pool.stats %>% select(Compound.Name,rsd) %>%
     mutate(`13C Acetyl CoA.Norm.Area` = rsd)

p <- ggplot(pool.stats, aes(x=rsd, text = Compound.Name)) + geom_histogram(binwidth = 0.05) + ggtitle("Relative Standard Deviation for 85 compounds in the pooled samples normalized to ...")
ggplotly(p)
```

##Compare normalizations
```{r}
pool.dat <- dat %>% filter(type == "Poo") %>% filter(ID.duplicate == "Poo-0614")
pool.dat <- pool.dat[-(grep("overload",x=pool.dat$Notes)),]

Norm.Names <- colnames(pool.dat)
Norm.Names

##Initial normalization
##13C Acetyl CoA.Norm.Area
pool.stats <-  pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`13C Acetyl CoA.Norm.Area`, na.rm=T), 
                sd = sd(`13C Acetyl CoA.Norm.Area`, na.rm=T), rsd = sd/m)
rsd.dat <- pool.stats %>% select(Compound.Name) %>%
     mutate(`13C Acetyl CoA.Norm.Area` = pool.stats$rsd)

##13C-Isethionic Acid.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`13C-Isethionic Acid.Norm.Area`, na.rm=T), 
               sd = sd(`13C-Isethionic Acid.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`13C-Isethionic Acid.Norm.Area` =
                                   pool.stats$rsd)
##13C-Sulfoacetic Acid.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`13C-Sulfoacetic Acid.Norm.Area`, na.rm=T), 
               sd = sd(`13C-Sulfoacetic Acid.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`13C-Sulfoacetic Acid.Norm.Area` =
                                   pool.stats$rsd)

##`13C-Sulfolactic Acid.Norm.Area`
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`13C-Sulfolactic Acid.Norm.Area`, na.rm=T), 
               sd = sd(`13C-Sulfolactic Acid.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`13C-Sulfolactic Acid.Norm.Area` =
                                   pool.stats$rsd)

##d3-Cysteic Acid.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`d3-Cysteic Acid.Norm.Area`, na.rm=T), 
               sd = sd(`d3-Cysteic Acid.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`d3-Cysteic Acid.Norm.Area` =
                                   pool.stats$rsd)

##d4 Succinic Acid.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`d4 Succinic Acid.Norm.Area`, na.rm=T), 
               sd = sd(`d4 Succinic Acid.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`d4 Succinic Acid.Norm.Area` =
                                   pool.stats$rsd)

##D-Methionine.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`D-Methionine.Norm.Area`, na.rm=T), 
               sd = sd(`D-Methionine.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`D-Methionine.Norm.Area` =
                                   pool.stats$rsd)

##Heavy Alanine.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`Heavy Alanine.Norm.Area`, na.rm=T), 
               sd = sd(`Heavy Alanine.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`Heavy Alanine.Norm.Area` =
                                   pool.stats$rsd)

##Heavy Arginine.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`Heavy Arginine.Norm.Area`, na.rm=T), 
               sd = sd(`Heavy Arginine.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`Heavy Arginine.Norm.Area` =
                                   pool.stats$rsd)

##Heavy Histidine.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`Heavy Histidine.Norm.Area`, na.rm=T), 
               sd = sd(`Heavy Histidine.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`Heavy Histidine.Norm.Area` =
                                   pool.stats$rsd)

##Heavy Isoleucine.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`Heavy Isoleucine.Norm.Area`, na.rm=T), 
               sd = sd(`Heavy Isoleucine.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`Heavy Isoleucine.Norm.Area` =
                                   pool.stats$rsd)

##Heavy Proline.Norm.Area
pool.stats <- pool.dat %>%
     group_by(Compound.Name) %>%
     summarise(m = mean(`Heavy Proline.Norm.Area`, na.rm=T), 
               sd = sd(`Heavy Proline.Norm.Area`, na.rm=T), 
               rsd = sd/m)
rsd.dat <- rsd.dat %>% mutate(`Heavy Proline.Norm.Area` =
                                   pool.stats$rsd)
```


##Normalize to optical density
Incomplete
```{r echo=FALSE, eval=FALSE}
dens <- read.xlsx("M:/Summer_2016_metabolites/Methods_Paper_Samples/Bacteria_Skyline_Samples/Cultures_Extraction_Log.xlsx", 1)  # read first sheet

dens.order <- order(dens$Vial.ID)[1:16]

new.df$OD <- as.numeric(as.character(dens$OD.1.10..at.Harvest.time[dens.order]))

OD.Norm.df <- apply(new.df,2,function(x) x/new.df[,"OD"])
```




